---
title: "ESA five-year reviews"
output:
  flexdashboard::flex_dashboard:
    css: custom.css
    social: menu
    df_print: paged
    navbar:
    - {title: "CCI", align: left, href: "https://cci-dev.org"}
    - {title: "Defenders", align: left, href: "http://www.defenders.org"}
    - {title: "", icon: "fa-question-circle fa-lg", align: right, href: "mailto:esa@defenders.org?subject=Five-year reviews app"}
    - {title: "", icon: "fa-github fa-lg", align: right, href: "https://github.com/Defenders-ESC/"}
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(dplyr)
library(DT)
library(ecosscraper)
library(flexdashboard)
library(ggplot2)
library(ggthemes)
library(plotly)
library(RPostgreSQL)
library(secret)
library(shiny)

load("TECP_table.rda")
TECP_names <- names(TECP_table)

con <- dbConnect(
  dbDriver("PostgreSQL"),
  dbname = "postgres",
  user = ESC_DB_USR,
  password = get_secret(ESC_DB_PW, key = priv_key, vault),
  host = ESC_DB
)

TECP_table <- dbSendQuery(con, "select * from tecp_table") %>%
  dbFetch()
names(TECP_table) <- TECP_names

load("fiveyr_table.rda")
load("counties_table_2016-12-17.rda")

THRES <- Sys.Date() - 365*5

listed <- filter(
  TECP_table, 
  Federal_Listing_Status != "Candidate" &
  Federal_Listing_Status != "Proposed Endangered" &
  Federal_Listing_Status != "Proposed Threatened" &
  Federal_Listing_Status != "Experimental Population, Non-Essential") %>%
  filter(U.S._or_ForeignListed != "Foreign")
listed$First_Listed <- as.Date(listed$First_Listed, format = "%b %d, %Y")

list_cnt <- left_join(listed, counties_table, by = "Scientific_Name")
fiveyr <- full_join(list_cnt, fiveyr_table, by = c("Scientific_Name" = "Species"))
fiveyr$Date <- as.Date(fiveyr$Date)
fiveyr <- filter(fiveyr, fiveyr$Date < as.Date("2017-01-31") | is.na(fiveyr$Date))

filter_state <- function(df, state) {
  if(state != "All") {
    tmp <- filter(df, State == state)
  } else {
    tmp <- df
  }
  tmp <- distinct(tmp, Scientific_Name, .keep_all = TRUE)
  tmp <- select(tmp, Common_Name, Scientific_Name, First_Listed,
                Species_Group, Federal_Listing_Status, Date, Title,
                Doc_Link)
  names(tmp) <- c("Common_Name",
                  "Scientific_Name",
                  "First_Listed",
                  "Taxon_Group",
                  "Listing_Status",
                  "Review_Date",
                  "Doc_Title",
                  "Doc_Link")
  return(tmp)
}

filter_taxon <- function(df, taxon) {
  if(taxon != "All") {
    return(filter(df, Taxon_Group == taxon))
  }
  return(df)
}

filter_need5yr <- function(df) {
  res <- filter(df, First_Listed < THRES)
  return(res)
}
```

Inputs {.sidebar}
-----------------------------------------------------------------------

```{r selector, echo = FALSE}
states <- sort(unique(list_cnt$State))
states <- c("All", states)
inputPanel(
  selectInput("sel_state",
              label = "State",
              choices = states,
              selected = "All")
)

taxon <- sort(unique(list_cnt$Species_Group))
taxon <- c("All", taxon)
inputPanel(
  selectInput("sel_taxon",
              label = "Taxonomic Group",
              choices = taxon,
              selected = "All")
)

cur_dat <- reactive(fiveyr %>%
                      filter_state(input$sel_state) %>%
                      filter_taxon(input$sel_taxon))
```

The [U.S. Endangered Species Act](https://www.fws.gov/endangered/laws-policies/)
requires a review of the status of listed species once every five years. Because
of budget and personnel constraints, these five-year reviews are often behind
schedule. This is a simple tool to view the timeliness of five-year reviews.

`r hr()`

```{r note, echo = FALSE}
helpText("Note: To ensure this app loads quickly, we use data collected on
17 Dec 2016 rather than real-time  scraping of the Fish and Wildlife Service's
website. We will update the data on a regular basis.")
```

`r hr()`

[Defenders of Wildlife](http://www.defenders.org)

[Contact us](mailto:esa@defenders.org)

Column {data-width=400}
-----------------------------------------------------------------------

### Timeliness

```{r timely, echo = FALSE}
df <- reactive({
  need_5yr <- filter_need5yr(cur_dat())
  n_ontime <- sum(need_5yr$Review_Date >= THRES, na.rm = TRUE)
  n_pastdue <- sum(need_5yr$Review_Date < THRES | is.na(need_5yr$Review_Date),
                   na.rm = TRUE)
  n_notdue <- dim(cur_dat())[1] - dim(need_5yr)[1]

  data.frame(category = c("On-time", "Past due", "Not due"),
             count = c(n_ontime, n_pastdue, n_notdue))
})

# reactive( saveRDS(df(), "~/Downloads/fiveyr_data.rds") )

t <- list(
  family = "'Open Sans'",
  size = 12,
  color = 'black'
)

renderPlotly({
  df() %>%
    plot_ly(
      x = df()$category,
      y = df()$count,
      type = "bar",
      marker = list(
        color = "#595959"
      )
    ) %>%
    layout(yaxis = list(title="# five-yr. reviews"))
})
```

> The timeliness of species' 5-year reviews as of Jan, 2017.

### Distribution of 5y Review Dates

```{r histogram, echo = FALSE}
# renderPlot({
#   ggplot(cur_dat(), aes(x = cur_dat()$Review_Date)) +
#     geom_histogram(bins = 12) +
#     geom_vline(xintercept = as.numeric(as.Date("2012-01-01")), colour = "red") +
#     labs(x = "Review Date",
#          y = "# of five-year reviews") +
#     theme_hc()
# })

renderPlotly({
  brks <- hist(cur_dat()$Review_Date, breaks = 12, plot = FALSE)
  # observe(print(brks))

  line <- list(
    type = "line",
    line = list(color = "red"),
    x0 = as.Date("2012-01-01"),
    x1 = as.Date("2012-01-01"),
    y0 = 0,
    y1 = max(brks$counts)
  )

  plot_ly(
    cur_dat(),
    x = cur_dat()$Review_Date,
    type = "histogram",
    marker = list(
      color = "#595959"
    ),
    nbinsx = 12
  ) %>%
  layout(yaxis = list(title="# five-yr. reviews"),
         shapes = line)
})
```

> The red line marks 01 Jan 2012, approx. five years ago.


Column {data-width=600}
-----------------------------------------------------------------------

### Data Table

```{r datatable, echo = FALSE}
DT::renderDataTable({
  minitab <- select(cur_dat(),
                    Scientific_Name, Common_Name,
                    First_Listed, Review_Date)
  DT::datatable(minitab,
                options = list(
                  rownames = FALSE,
                  pageLength = 10,
                  lengthMenu = c(10, 12, 15, 20))
                )
})
```

> "Review_Date" is the date of the most recent 5-year review on [ECOS](http://ecos.fws.gov). If the "First_Listed" date is less than 5 years ago, then a species does not need a 5-year review. A missing "Review_Date" means there is no 5-year review on ECOS.
